---
title: 'Bradford road traffic accidents'
author: "Dobgima Mofor"
#date: "2023-11-29"
output:
  pdf_document:
    toc: yes
  html_document:
    code_folding: hide
    number_sections: yes
    self_contained: yes
    toc: yes
    toc_float: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r, results='hide', message=FALSE, warning=FALSE, echo=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo=FALSE) 

library(tidyverse)
library(flextable)
library(forcats)
library(lubridate)
library(here)

options(shiny.usecairo=T)
setalloccol = data.table::setalloccol
```

## Road traffic accidents in Bradford

**Task** The Director of Neighbourhoods wants to understand more about
road traffic crashes in Bradford. You, the HDRC data scientist, have
been asked to give some insight into road traffic crashes in Bradford.

*Note* Availability of more specific dates rather than years would allow
for calculation of rates

Datasets

```{r}
list.files()
```

```{r}
casualties <- read_csv("bfd_casualties.csv", show_col_types = FALSE)
crashes <- read.csv("bfd_crashes.csv")
vehicles <- read.csv("bfd_vehicles.csv")

attach(casualties)
```

#1. Descriptives

```{r}
casualty_links <- casualties %>% 
  left_join(crashes, by = c('accident_index','accident_reference')) %>% 
                          mutate(crash_date = date(datetime)) %>% 
                          mutate(crash_month = month(datetime, label = TRUE))

```


```{r, warning=FALSE}
nrow(distinct(casualties %>% select(accident_index)))

nrow(casualties)
range(casualty_links$datetime)
```

There were 2,518 accidents from 2018 to 2020 which resulted in 3,350 casualties. Most casualties did not result in deaths (85% Slight and 14% Serious), and overall, casualties decreased across all severity categories from 2018 to 2020.


```{r, warning=FALSE}
round(chisq.test(table(casualties$accident_year,casualties$sex_of_casualty))$p.value,3)
round(chisq.test(table(casualties$accident_year, casualties$age_band_of_casualty))$p.value,3)
```
# Table for accident rates
```{r}
accident_links <- casualties %>% 
  filter(casualty_reference == 1) %>%                           #select unique accident
  left_join(crashes, by = c('accident_index','accident_reference')) %>% 
                          mutate(crash_date = date(datetime)) %>% 
                          mutate(crash_month = month(datetime, label = TRUE))

accident_month_2018 <- accident_links %>%
  filter(accident_year.x == 2018) %>% 
  group_by(crash_month) %>% 
  summarise(`2018` = n()) %>% 
  ungroup() %>% 
  arrange(desc(`2018`))

accident_month_2019 <- accident_links %>%
  filter(accident_year.x == 2019) %>% 
  group_by(crash_month) %>% 
  summarise(`2019` = n()) %>% 
  ungroup() %>% 
  arrange(desc(`2019`))

accident_month_2020 <- accident_links %>%
  filter(accident_year.x == 2020) %>% 
  group_by(crash_month) %>% 
  summarise(`2020` = n()) %>% 
  ungroup() %>% 
  arrange(desc(`2020`))

accident_month <- accident_month_2018 %>% left_join(accident_month_2019, by = ('crash_month')) %>% 
    left_join(accident_month_2020, by = ('crash_month')) %>% 
  pivot_longer(!crash_month, names_to = 'year', values_to = 'accidents')

accident_month
```


# Table for casualty rates
```{r}
casualty_month_2018 <- casualty_links %>%
  filter(accident_year == 2018) %>% 
  group_by(crash_month) %>% 
  summarise(`2018` = n()) %>% 
  ungroup() %>% 
  arrange(desc(`2018`))

casualty_month_2019 <- casualty_links %>%
  filter(accident_year == 2019) %>% 
  group_by(crash_month) %>% 
  summarise(`2019` = n()) %>% 
  ungroup() %>% 
  arrange(desc(`2019`))

casualty_month_2020 <- casualty_links %>%
  filter(accident_year.x == 2020) %>% 
  group_by(crash_month) %>% 
  summarise(`2020` = n()) %>% 
  ungroup() %>% 
  arrange(desc(`2020`))

casualty_month <- casualty_month_2018 %>% left_join(casualty_month_2019, by = ('crash_month')) %>% 
    left_join(casualty_month_2020, by = ('crash_month')) %>% 
  pivot_longer(!crash_month, names_to = 'year', values_to = 'casualties')

casualty_month


acc_cas <- accident_month %>%  left_join(casualty_month, by = c('crash_month', 'year'))
```


# Accident and casualty rate plot
```{r}
p1 <- ggplot(data = acc_cas, #=dose, y=len, group=sup
       aes(x = crash_month, col = year)) +
#  facet(vars(year)) +
geom_line(aes(y = casualties, group = year, linetype= 'Casualties'), linewidth=0.7) +
geom_line(aes(y = accidents, group = year, linetype= 'Accidents'), linewidth=0.7) +
    labs(title = 'Rate of accidents in Bradford 2018-2020',
      x = 'Time (months)',
      y = 'Number of monthly casualties'
  ) +
  theme_bw() +
  theme(axis.text = element_text(size = 10, vjust = 0.5, hjust=0.5),
        axis.text.x = element_text(margin = margin(t = 0, unit = "cm")))
```
Accidents and casualties were lowest in the month of April, and peaked from the September to December months. The fewest cassualties were seen in the month of April

```{r}
ggsave('Casualty and accident rates.png',p1, path = here("plots"), width = 7, height = 7)

ggsave('Casualty and accident rates.png',p1, path = here("pngs"), width = 7, height = 7)

```


```{r}
ggplot(data = casualty_month, #=dose, y=len, group=sup
       aes(x = crash_month, y = casualties, group = year, col = year)) +
#  facet_grid(vars(year)) +
geom_line() +
    labs(title = 'Rate of casualties in Bradford 2018-2020',
      x = 'Time (months)',
      y = 'Number of casualties per month'
  ) +
  theme_bw() +
  theme(axis.text = element_text(size = 10, vjust = 0.5, hjust=0.5),
        axis.text.x = element_text(margin = margin(t = 0, unit = "cm")))
```

# Casualty rate plot

```{r}

casualty_month_Fatal <- casualty_links %>%
  filter(casualty_severity == 'Fatal') %>% 
  group_by(crash_month) %>% 
  summarise(Fatal = n()) %>% 
  ungroup() %>% 
  arrange(desc(Fatal))

casualty_month_Serious <- casualty_links %>%
  filter(casualty_severity == 'Serious') %>% 
  group_by(crash_month) %>% 
  summarise(Serious = n()) %>% 
  ungroup() %>% 
  arrange(desc(Serious))

casualty_month_Slight <- casualty_links %>%
  filter(casualty_severity == 'Slight') %>% 
  group_by(crash_month) %>% 
  summarise(Slight = n()) %>% 
  ungroup() %>% 
  arrange(desc(Slight))

casualty_severity <- casualty_month_Fatal %>% left_join(casualty_month_Serious, by = ('crash_month')) %>% 
    left_join(casualty_month_Slight, by = ('crash_month')) %>% 
  pivot_longer(!crash_month, names_to = 'severity', values_to = 'casualties')


write.table(casualty_severity, file = "tables/ casualty severity.csv",
            sep = ",", row.names = F, col.names = TRUE)

casualty_severity

```

```{r}
p2 <- ggplot(data = casualty_severity, 
       aes(x = crash_month, col = severity)) +
#  facet(vars(year)) +
geom_line(aes(y = casualties, group = severity, linetype= 'Casualties'), linewidth=0.7) +
#geom_line(aes(y = accidents, group = year, linetype= 'Accidents'), linewidth=0.7) +
    labs(title = 'Rate of accidents in Bradford 2018-2020',
         subtitle = 'Stratified by severity',
      x = 'Time (months)',
      y = 'Number of monthly casualties'
  ) +
  theme_bw() +
  theme(axis.text = element_text(size = 10, vjust = 0.5, hjust=0.5),
        axis.text.x = element_text(margin = margin(t = 0, unit = "cm")))

ggsave('Casualty severity rate.png',p2, path = here("plots"), width = 7, height = 7)

ggsave('Casualty severity rate.png',p2, path = here("pngs"), width = 7, height = 7)

p2
```


```{r}
df_sev <- table(casualties$casualty_severity,casualties$accident_year)

df <- data.frame(cbind(df_sev[,1],df_sev[,2],df_sev[,3]))
names(df) <- c('2018','2019',"2020")
df <- df %>% mutate(`2018` = ((`2020` - `2018`)/`2020`)*100) %>% 
  mutate(`2018` = round(`2018`,1)) %>% 
  mutate(`2019` = ((`2020` - `2019`)/`2020`)*100) %>% 
  mutate(`2019` = round(`2019`,1))

df_severity <- df %>% select('2020','% change from 2018'= '2018', '% change from 2019'='2019')

df_severity

write.table(df_severity, file = "tables/ casualty severity changes.csv",
            sep = ",", row.names = F, col.names = TRUE)
```

## Casualty class

Drivers were more than twice as likely to be involved in accidents as
Pedestrians, even after adjusting for sex

```{r}
p3 <- ggplot(data = casualties,
       aes(x = fct_infreq(casualty_class)), y = no_acc) +
  geom_bar(fill = '#44846fff') +
  facet_wrap(vars(sex_of_casualty)) +
  labs(title = 'Number of accident casualties by casualty class',
       subtitle = 'Stratified by year of accident',
       x = 'Casualty class',
       y = 'Number of accidents') +
  theme_bw() 

ggsave('Casualty class by accident year.pdf',p3, path = here("plots"), width = 7, height = 7)
ggsave('Casualty class by accident year.png',p3, path = here("pngs"), width = 7, height = 7)

```

## Casualties by road user type

Car occupants, pedestrians and cyclists were the most accident (200+ each)
casualties compared to other road users

**Table 1: Reported road casualty counts by road user type in Bradford,
2018,2019 and 2020**

```{r}
temp <- casualties %>%
# Remove missing data
  mutate(tag = ifelse(age_band_of_casualty %in% 'Data missing or out of range',1,0)) %>%
  slice(-(which(tag == 1))) %>% 
  select(-tag) %>% 
  
  group_by(casualty_type) %>% 
  mutate(cas_seq = row_number()) %>%
  mutate(tot_cas = max(cas_seq)) %>% 
  ungroup() %>% 
  arrange(casualty_type) %>% 
  mutate(tag = 0) %>% 
  mutate(tag = ifelse(tot_cas >= 200, 1,tag)) %>% 
#  mutate(tag = ifelse(tag == 0 & tot_cas >= 20,2,tag)) %>% 
#  mutate(tag = ifelse(tag == 0 & tot_cas< 20,3,tag)) %>%
  mutate(accident_year = factor(accident_year))

temp1 <- temp %>% 
  filter(cas_seq == 1) %>% 
  arrange(desc(tot_cas)) %>% 
  select(casualty_type, casualty_count = tot_cas) %>% 
  rename(`Casualty type` = "casualty_type", `Casualty count` = "casualty_count")

 flextable(temp1, cwidth = 2, col_keys = names(temp1))

write.table(temp1, file = "tables/ Casualty_type_count.csv",
            sep = ",", row.names = F, col.names = TRUE)
```

```{r}
df_typ <- table(casualties$casualty_type,casualties$accident_year)
```

**Table 2: Changes in casualty counts by road user type in Bradford,
2018,2019 and 2020**

```{r}
df1 <- data.frame(cbind(df_typ[,1],df_typ[,2],df_typ[,3]))
names(df1) <- c('2018','2019',"2020")
df1 <- df1 %>% mutate(`2018` = ((`2020` - `2018`)/`2020`)*100) %>% 
  mutate(`2018` = round(`2018`,1)) %>% 
  mutate(`2019` = ((`2020` - `2019`)/`2020`)*100) %>% 
  mutate(`2019` = round(`2019`, 1))

df_type <- df1 %>% cbind(rownames(df1)) %>% 
  select('Class'='rownames(df1)' ,'2020','% change from 2018'= '2018', '% change from 2019'='2019') %>% 
  arrange(desc(`2020`))

flextable(df_type, cwidth = 1)

write.table(df_type, file = "tables/ Changes in casualty by vehicle type.csv",
            sep = ",", row.names = F, col.names = TRUE)
```

Among the most at risk road users (200+ total casualties), the most 
improvement was seen in 2020 was from car occupants (casualties decreased by 
43% compared to 2018 and 23.8% compared to 2019). Conversely, there were 
more cyclist casualties in 2020 than the preceding year (9.1% increase from 2019).
Of the remaining road users who witnessed casualty increments, 5 involved either 
a motorcycle or mobility scooter.

### 200+ Casualties
```{r}
# Remove missing data
temp_imd <-  temp %>% 
  mutate(tag = ifelse(casualty_imd_decile %in% 'Data missing or out of range',1,0)) %>%
  slice(-(which(tag == 1))) %>% 
  select(-tag)
```

```{r}

# Remove missing data
temp_imd <-  temp %>% 
  mutate(tag = ifelse(casualty_imd_decile %in% 'Data missing or out of range',1,0)) %>%
  slice(-(which(tag == 1))) %>% 
  select(-tag) 

my_colors <- c("#006853ff", "#ff7733ff", '#001a70ff')  
names(my_colors) <- levels(factor(levels(temp$accident_year)))
my_scale <- scale_fill_manual(name = "Year", values = my_colors) 


p4 <- ggplot(data = temp_imd %>% 
         filter(),
       aes(x = fct_infreq(casualty_imd_decile), fill = fct_infreq(accident_year))) +
 #   facet_wrap(vars(sex_of_casualty)) +
  geom_bar(position="dodge") +
  labs(title = 'Reported road casualties by index of multiple deprivation in Bradford, 2020 compared with 2018',
         subtitle = '200+ total casualties',
      x = 'Level of deprivation',
      y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text = element_text(size = 10, angle = 90, vjust = 0.5, hjust=0.5),
        axis.text.x = element_text(margin = margin(t = 0, unit = "cm"))) +
  my_scale

ggsave('Road casualties by IMD.pdf', p4, path = here("plots"), width = 7, height = 7)

ggsave('Road casualties by IMD.png', p4, path = here("pngs"), width = 7, height = 7)
```

Road casualties were directly proportional to the level of deprivation.
The most deprived areas witnessed the most casualties across all years.


# Casualty by age category, adjusting for year of casualty

```{r}
my_colors <- c("#006853ff", "#ff7733ff",'#001a70ff')  
names(my_colors) <- levels(factor(levels(temp$accident_year)))
my_scale <- scale_fill_manual(name = "Sex", values = my_colors) 


p5 <- ggplot(data = temp %>% 
         filter(),
       aes(x = age_band_of_casualty, fill = fct_infreq(accident_year))) +
#    facet_wrap(vars(sex_of_casualty)) +
  geom_bar(position="dodge") +
  labs(title = 'Reported road casualties by index of multiple deprivation in Bradford, 2020 compared with 2018',
         subtitle = '200+ total casualties',
      x = 'Age band of casualties',
      y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text = element_text(size = 10, angle = 90, vjust = 0.5, hjust=0.5),
        axis.text.x = element_text(margin = margin(t = 0, unit = "cm"))) +
  my_scale

p5

ggsave('Road casualties by IMD.pdf', p5, path = here('plots'), width = 7, height = 7)

ggsave('Road casualties by IMD.png', p5, path = here('pngs'), width = 7, height = 7)
```

Majority of accident casualties were in their active years (16-65 years)

```{r}
a1<- casualties %>%
  filter(accident_year == 2018) %>% 
 group_by(age_band_of_casualty) %>% 
summarise(A2018 = n())

a2<- casualties %>%
  filter(accident_year == 2019) %>% 
 group_by(age_band_of_casualty) %>% 
summarise(A2019 = n())

a3<- casualties %>%
  filter(accident_year == 2020) %>% 
 group_by(age_band_of_casualty) %>% 
summarise(A2020 = n())

Casualty_age <- left_join(a1,a2, by = "age_band_of_casualty") %>% 
  left_join(a3, by = "age_band_of_casualty") %>% 
  mutate(A2019 = (A2020-A2019)/A2020) %>% 
  mutate(A2018 = (A2020-A2018)/A2020) %>% 
  mutate(A2018 = round(A2018*100,2)) %>% 
  mutate(A2019 = round(A2019*100,2)) %>% 
  rename(Age = age_band_of_casualty, `% change from 2018` = A2018, `% change from 2019` = A2019, `2020` = A2020) %>% 
  select(Age,`2020`,`% change from 2019`,`% change from 2018`)

write.table(Casualty_age, file = "tables/ Changes in casualty by age group.csv",
            sep = ",", row.names = F, col.names = TRUE)

flextable(Casualty_age, cwidth = 1)
```



```{r}
df_age_m <- casualties %>% filter(sex_of_casualty == 'Male') %>% select(age_band_of_casualty,accident_year) %>% table()
df_age_f <- casualties %>% filter(sex_of_casualty == 'Female') %>% select(age_band_of_casualty,accident_year) %>% table()
```

Table 4: Reported road casualties by road user type in Great Britain,
2021, compared with 2020, 2019 and 2011

```{r}
df2m <- data.frame(cbind(df_age_m[,1],df_age_m[,2], df_age_m[,3])) %>% mutate(Sex = 'Male')
df2f <- data.frame(cbind(df_age_f[,1],df_age_f[,2], df_age_f[,3])) %>% mutate(Sex = 'Female')
df2 <- rbind(df2m, df2f)
names(df2) <- c('2018','2019','2020', 'Sex')
df2 <- df2 %>% mutate(`2018` = ((`2020` - `2018`)/`2020`)*100) %>% 
  mutate(`2018` = round(`2018`,1)) %>% 
  mutate(`2019` = ((`2020` - `2019`)/`2020`)*100) %>% 
  mutate(`2019` = round(`2019`, 1))

df_sex <- df2 %>% 
  select('Sex','2020','% change from 2018'= '2018', '% change from 2019'='2019') %>% 
  arrange(desc(Sex))

Age <- c("0 - 5", "11 - 15", "16 - 20", "21 - 25", "26 - 35", "36 - 45", "46 - 55", "56 - 65", "6 - 10", "66 - 75", "Data missing or out of range", "Over 75", "0 - 5", "11 - 15", "16 - 20", "21 - 25", "26 - 35", "36 - 45", "46 - 55", "56 - 65", "6 - 10", "66 - 75", "Data missing or out of range", "Over 75")


df_sex1 <- factor(Age)  %>%
  cbind(df_sex) %>%
  group_by(Sex) %>% 
  arrange(Age) %>%
  ungroup()%>%
  arrange(desc(Sex)) %>%data.frame()

df_sexa <- df_sex1 %>% 
  slice(-11,-23) %>%
  group_by(Sex) %>% 
  mutate(ind = row_number()) %>% 
  ungroup() %>% 
  mutate(tag = 2)

# Re-order the rows

for (i in 2:nrow(df_sexa)) {
  if(df_sexa$ind[i] > 1 & df_sexa$ind[i] != 9)
    {
    df_sexa$tag[i] <- df_sexa$ind[i] + 1
  }
}

df_sexa$tag[df_sexa$ind == 1] <- 1

df_sexa <- df_sexa %>% 
                arrange(desc(Sex), tag) %>% 
                select(-ind, -tag) %>% 
                data_frame()

df_sexa <- df_sexa %>%
  data.frame() %>% 
  as_grouped_data(groups = "Sex")

names(df_sexa) <- c('Age cat','Sex', '2020', '% change from 2018', '% change from 2019')

write.table(df_sexa, file = "tables/ Changes in casualty by age and sex type.csv",
            sep = ",", row.names = F, col.names = TRUE)

df_sex <- flextable(df_sexa, names(df_sexa), cwidth = 1)
df_sex
```

For males, an increase in casualties was seen in 2020 among 
children (6 - 10 years) and young-middle aged adults (21-25 and 26-35 years). 
Similarly, female casualties increased among infants (0-5 years, 9.1% from 2018) 
and middle adults (26-35 years). The most improvement was in road users 
over 75 years for females (-166.7% form 2018) and 46-55 years for males (-115.2% from 2019).

### Examine age categories with most casualties

The casualty increments in 2020 were in the age categories 0-5 years, 6-10 years, 16-20, 21-25 years and 26-35 years.

```{r}
temp$sex_of_casualty <- temp$sex_of_casualty %>% factor(levels = c('Male','Female'))
temp$age_band_of_casualty <- temp$age_band_of_casualty %>% factor()


#(levels = c( '0 - 5', '11 - 15', '16 - 20', '21 - 25', '26 - 35', '36 - 45', '46 - 55',  #'56 - 65', '6 - 10', '66 - 75', 'Over 75 '))

temp2 <- temp %>% 
  mutate(tag2 = 0) %>% 
  mutate(tag2 = ifelse(age_band_of_casualty %in% '0 - 5','0 to 10 years',tag2)) %>% 
  mutate(tag2 = ifelse(tag2 == 0 & age_band_of_casualty %in% '6 - 10','0 to 10 years',tag2)) %>% 
  mutate(tag2 = ifelse(tag2 == 0 & age_band_of_casualty %in% '21 - 25','21 to 35 years',tag2)) %>%
  mutate(tag2 = ifelse(tag2 == 0 & age_band_of_casualty %in% '26 - 35','21 to 35 years',tag2))
  
temp2 <- temp %>% 
  mutate(tag2 = 0) %>% 
  mutate(tag2 = ifelse(age_band_of_casualty %in% '0 - 5','0 - 5',tag2)) %>% 
  mutate(tag2 = ifelse(tag2 == 0 & age_band_of_casualty %in% '6 - 10','6 - 10',tag2)) %>% 
  mutate(tag2 = ifelse(tag2 == 0 & age_band_of_casualty %in% '16 - 20','16 - 20',tag2)) %>%
  mutate(tag2 = ifelse(tag2 == 0 & age_band_of_casualty %in% '21 - 25','21 - 25',tag2)) %>%
  mutate(tag2 = ifelse(tag2 == 0 & age_band_of_casualty %in% '26 - 35','26 - 35',tag2))


casualty_cat_most <- temp2 %>% 
  filter(tag2 != 0) %>%
  mutate(tag = ifelse(tag == 1, '200 or more casualties','Less than 200 casualties')) %>%
  filter(tag == '200 or more casualties')

table(casualty_cat_most$tag2,casualty_cat_most$sex_of_casualty)
```


```{r}

my_colors1 <- c("#006853ff", '#001a70ff')  
names(my_colors) <- levels(factor(levels(temp$sex_of_casualty)))
my_scale1 <- scale_fill_manual(name = "Sex", values = my_colors) 


p6 <- ggplot(data = casualty_cat_most %>% 
               filter(tag2 == '0 - 5' & sex_of_casualty == 'Female'),
       aes(x = fct_infreq(casualty_type), fill = sex_of_casualty)) +
  geom_bar(position="dodge") +
  facet_wrap(vars(accident_year)) +
  labs(title = 'Reported road casualties by road user type in Bradford, 2018-2020',
       subtitle = 'Children (0-5 years) involved in accident where casualty class recorded 200 or more casualties',
       x = 'Casualty class',
       y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale1

p6

ggsave('Plot of Females 0-6 years.pdf', p6, path = here('plots'), width = 7, height = 7)

ggsave('Plot of Females 0-6 years.png', p6, path = here('pngs'), width = 7, height = 7)
```

Although females casualties increased across the 3 years, there were below 8 casualties per year. Car occupant casualties increased each year. In the 6-10 year category, pedestrians had more than double the number of casualties seen in cyclists or car occupants.

```{r}
p7 <- ggplot(data = casualty_cat_most %>% filter(tag2 == '6 - 10' & sex_of_casualty == 'Male'),
       aes(x = fct_infreq(casualty_type), fill = sex_of_casualty)) +
  geom_bar(position="dodge") +
  facet_wrap(vars(accident_year)) +
  labs(title = 'Reported road casualties by road user type in Bradford, 2018-2020',
       subtitle = 'Children (6-10) involved in accident where casualty class recorded 200 or more casualties',
       x = 'Casualty class',
       y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale1

p7

ggsave('Plot of Females 6-10 years.pdf', p7, path = here('plots'), width = 7, height = 7)

ggsave('Plot of Females 6-10 years.png', p7, path = here('pngs'), width = 7, height = 7)
```



```{r}
p7 <- ggplot(data = casualty_cat_most %>% filter(tag2 == '16 - 20' & sex_of_casualty == 'Male'),
       aes(x = fct_infreq(casualty_type), fill = sex_of_casualty)) +
  geom_bar(position="dodge") +
  facet_wrap(vars(accident_year)) +
  labs(title = 'Reported road casualties by road user type in Bradford, 2018-2020',
       subtitle = 'Children (16-20) involved in accident where casualty class recorded 200 or more casualties',
       x = 'Casualty class',
       y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale1

p7

ggsave('Plot of Females 16-20 years.png', p6, path = here('plots'), width = 7, height = 7)
```



In age categories with over 100 casualties per gender, car occupants were more
than 7 times involved in casualties than cyclists or pedestrians

```{r}
p8 <- ggplot(data = casualty_cat_most %>% filter(tag2 == '21 - 25' & sex_of_casualty == 'Male'),
       aes(x = fct_infreq(casualty_type), fill = sex_of_casualty)) +
  geom_bar(position="dodge") +
  facet_wrap(vars(accident_year)) +
  labs(title = 'Reported road casualties by road user type in Bradford, 2018-2020',
       subtitle = 'Children (21-25) involved in accident where casualty class recorded 200 or more casualties',
       x = 'Casualty class',
       y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale1

p8

ggsave('Plot of Females 21-25 years.pdf', p8, path = here('plots'), width = 7, height = 7)

ggsave('Plot of Females 21-25 years.png', p8, path = here('pngs'), width = 7, height = 7)
```

```{r}
p9 <- ggplot(data = casualty_cat_most %>% filter(tag2 == '26 - 35'),
       aes(x = fct_infreq(casualty_type), fill = sex_of_casualty)) +
  geom_bar(position="dodge") +
  facet_wrap(vars(accident_year)) +
  labs(title = 'Reported road casualties by road user type in Bradford, 2018-2020',
       subtitle = 'Adults (26-35) involved in accident where casualty class recorded 200 or more casualties',
       x = 'Casualty class',
       y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale1


p9

ggsave('Plot of Adults 26-36 years.pdf', p9, path = here('plots'), width = 7, height = 7)

ggsave('Plot of Adults 26-36 years.png', p9, path = here('pngs'), width = 7, height = 7)
```

```{r}
p10 <- ggplot(data = casualty_cat_most,
       aes(x = fct_infreq(tag2))) +
  geom_bar(position="dodge", fill = 'blue') +
  facet_wrap(vars(casualty_type)) +
  labs(title = 'Reported road casualties by age category stratified by casualty type in Bradford, 2018-2020',
       subtitle = '',
       x = 'Casualty class',
       y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale1

p10

ggsave('Plot of road categories for at risk group by age category and casualty type.pdf', p10, path = here('plots'), width = 7, height = 7)

ggsave('Plot of road categories for at risk group by age category and casualty type.png', p10, path = here('pngs'), width = 7, height = 7)

```


```{r}
p11 <- ggplot(data = casualty_cat_most,
       aes(x = fct_infreq(tag2))) +
  geom_bar(position="dodge", fill = 'blue') +
  facet_wrap(vars(casualty_class)) +
  labs(title = 'Reported road casualties of the top 3 casualty types by age category type in Bradford, 2018-2020',
       subtitle = 'Stratified by casualty class',
       x = 'Casualty age group',
       y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

p11

ggsave('Plot of road casualties for top 3 casualty types stratified by age.pdf', p11, path = here('plots'), width = 7, height = 7)
ggsave('Plot of road casualties for top 3 casualty types stratified by age.png', p11, path = here('pngs'), width = 7, height = 7)

```


```{r}
crashes <- crashes %>% mutate(accident_year = factor(accident_year))

casualty_young <- casualty_cat_most %>% 
  left_join(crashes, by = c('accident_index', 'accident_year', 'accident_reference')) %>% 
  filter(casualty_type == 'Car occupant')
```

Examine car
```{r}
table(casualty_young$bus_or_coach_passenger)

table(casualty_young$car_passenger)

table(casualty_young$age_band_of_casualty, casualty_young$casualty_class)

```

```{r}
p12 <- ggplot(data = casualty_young,
       aes(x = fct_infreq(casualty_class), fill = sex_of_casualty)) +
  geom_bar(position="dodge") +
  facet_wrap(vars(car_passenger)) +
  labs(title = 'Reported road casualties by road user type in Bradford, 2020',
       subtitle = 'Adults (26-35) involved in accident where casualty class recorded 200 or more casualties in 2020',
       x = 'Casualty class',
       y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale1

p12

ggsave('Plot of road casualties by road user type stratified by casualty class.pdf', p12, path = here('plots'), width = 7, height = 7)

ggsave('Plot of road casualties by road user type stratified by casualty class.png', p12, path = here('pngs'), width = 7, height = 7)

```

None were bus or coach accidents

```{r}
c1<- casualty_young %>% 
  filter(accident_year == '2020') %>% 
  group_by(light_conditions) %>%
  select(casualty_type, light_conditions) %>% 
  summarise(`2020` = n()) %>% 
  ungroup()

c2 <- casualty_young %>% 
  filter(accident_year == '2019') %>% 
  group_by(light_conditions) %>%
  select(casualty_type, light_conditions) %>% 
  summarise(`2019` = n()) %>% 
  ungroup() 

c3 <- casualty_young %>% 
  filter(accident_year == '2018') %>% 
  group_by(light_conditions) %>%
  select(casualty_type, light_conditions) %>% 
  summarise(`2018` = n()) %>% 
  ungroup() 

light_conditions <- left_join(c1,c2, by = 'light_conditions') %>% left_join(c3, by = 'light_conditions')

df_light <- light_conditions %>% arrange(desc(`2020`))

# chisq.test(light_conditions$`2020`,light_conditions$`2018`)

write.table(df_light, file = "tables/Environmental conditions/ Light type.csv",
            sep = ",", row.names = F, col.names = TRUE)

```


```{r}
#casualty_young %>% 
#  group_by(weather_conditions) %>%
#  select(casualty_type, weather_conditions) %>% 
#  summarise(no_of_casualties = n()) %>% 
#  ungroup() %>% 
#  arrange(desc(no_of_casualties))
```
# Weather conditions

```{r}
c1<- casualty_young %>% 
  filter(accident_year == '2020') %>% 
  group_by(weather_conditions) %>%
  select(casualty_type, weather_conditions) %>% 
  summarise(`2020` = n()) %>% 
  ungroup()

c2 <- casualty_young %>% 
  filter(accident_year == '2019') %>% 
  group_by(weather_conditions) %>%
  select(casualty_type, weather_conditions) %>% 
  summarise(`2019` = n()) %>% 
  ungroup() 

c3 <- casualty_young %>% 
  filter(accident_year == '2018') %>% 
  group_by(weather_conditions) %>%
  select(casualty_type, weather_conditions) %>% 
  summarise(`2018` = n()) %>% 
  ungroup() 

weather_conditions <- left_join(c1,c2, by = 'weather_conditions') %>% left_join(c3, by = 'weather_conditions')

weather <- weather_conditions %>% arrange(desc(`2020`))

write.table(df_light, file = "tables/Environmental conditions/ weather.csv",
            sep = ",", row.names = F, col.names = TRUE)

```

Most casualties among children occurred during the day, with lights and
in fine weather condition.

# Road surface conditions

```{r}
c1<- casualty_young %>% 
  filter(accident_year == '2020') %>% 
  group_by(road_type) %>%
  select(casualty_type, road_type) %>% 
  summarise(`2020` = n()) %>% 
  ungroup()

c2 <- casualty_young %>% 
  filter(accident_year == '2019') %>% 
  group_by(road_type) %>%
  select(casualty_type, road_type) %>% 
  summarise(`2019` = n()) %>% 
  ungroup() 

c3 <- casualty_young %>% 
  filter(accident_year == '2018') %>% 
  group_by(road_type) %>%
  select(casualty_type, road_type) %>% 
  summarise(`2018` = n()) %>% 
  ungroup() 

road_type <- left_join(c1,c2, by = 'road_type') %>% left_join(c3, by = 'road_type')

df_road <- road_type %>% arrange(desc(`2020`))

write.table(df_road, file = "tables/Environmental conditions/ road.csv",
            sep = ",", row.names = F, col.names = TRUE)

```

```{r}
#casualty_young %>% 
#  group_by(road_type) %>%
#  select(casualty_type, road_type) %>% 
#  summarise(no_of_casualties = n()) %>% 
#  ungroup() %>% 
#  arrange(desc(no_of_casualties))

```

```{r}
casualty_young %>% 
  group_by(urban_or_rural_area) %>%
  select(casualty_type, urban_or_rural_area) %>% 
  summarise(no_of_casualties = n()) %>% 
  ungroup() %>% 
  arrange(desc(no_of_casualties))


```

```{r}
casualty_young_fine <- casualty_young %>% 
  filter(weather_conditions == 'Fine no high winds') %>% 
  filter(light_conditions == 'Daylight') %>% 
  filter(road_surface_conditions == 'Dry') %>% 
  filter(road_type == 'Single carriageway') %>% 
  mutate(fine_tag = 1)

good_conditions <- nrow(casualty_young_fine)/nrow(casualty_young) 
good_conditions <- round(good_conditions *100,2)

cat(good_conditions, '% of casualties occured under good weather and road conditions')

```


```{r, warning=FALSE}
casualty_young_cause <- casualty_young %>% 
  left_join(casualty_young_fine %>% 
              select(accident_index, casualty_reference,accident_reference, fine_tag), by = c('accident_index', 'casualty_reference', 'accident_reference')) %>%
  filter(is.na(fine_tag)) %>%
  select(-fine_tag)

less_optimal_conditions <- nrow(casualty_young_cause)/nrow(casualty_young) 
less_optimal_conditions <- round(less_optimal_conditions *100,2)


cat(less_optimal_conditions, '% of casualties occured under less optimal weather and road conditions')

```

```{r}
p13 <- ggplot(data = casualty_young_fine %>% filter(
  #pedestrian_crossing_physical_facilities == 'No physical crossing facilities within 50 metres'
  ),
       aes(x = fct_infreq(pedestrian_crossing_physical_facilities))) +
  geom_bar(position="dodge", fill = 'blue') +
  facet_wrap(vars(accident_year)) +
  labs(title = 'Reported road casualties by accident year in Bradford, 2018-2020',
       subtitle = 'Available pedestrian crossing physical facilities',
       x = 'Pedestrian crossing physical facilities',
       y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale

p13

ggsave('Road casualties by accident year, adjusting for pedestrian crossings.pdf', p13, path = here('plots'), width = 7, height = 7)

ggsave('Road casualties by accident year, adjusting for pedestrian crossings.png', p13, path = here('pngs'), width = 7, height = 7)

```

```{r}
p14 <- ggplot(data = casualty_young_fine,
       aes(x = fct_infreq(age_band_of_casualty))) +
  geom_bar(position="dodge", fill = 'blue3') +
  facet_wrap(vars(pedestrian_crossing_human_control)) +
  labs(title = 'Reported road casualties by user age band in Bradford, 2018-2020',
       subtitle = 'Presence of a pedestrian crossing or human control',
       x = 'Casualty age band',
       y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale

p14

ggsave('Road casualties by age category, adjusting for pedestrian crossings human control.pdf', p14, path = here('plots'), width = 7, height = 7)

ggsave('Road casualties by age category, adjusting for pedestrian crossings human control.png', p14, path = here('pngs'), width = 7, height = 7)

```

```{r}
casualty_young_fine_v <- casualty_young_fine %>%
  left_join(vehicles, by = c('accident_index','accident_reference')) #%>% 
#  mutate(accident_year = accident_year.x)

```


```{r}
p15 <- ggplot(data = casualty_young_fine_v %>% 
         filter(pedestrian_crossing_physical_facilities != 'No physical crossing facilities within 50 metres' & junction_control != 'Give way or uncontrolled' & junction_control != 'Data missing or out of range'),
       aes(x = vehicle_manoeuvre)) +
  geom_bar(position="dodge", fill = 'blue') +
 # facet_wrap(vars(vehicle_left_hand_drive)) +
  labs(title = 'Reported road casualties by road user type in Bradford, 2018-2020',
       subtitle = 'In the presence of some physical crossing facilities and with controlled junctions',
       x = 'Vehicle manoeuvre',
       y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale

ggsave('Road casualties by vehicle manoevre, in the presence of pedestrian crossings and human control.pdf', p15, path = here('plots'), width = 7, height = 7)

ggsave('Road casualties by vehicle manoevre, in the presence of pedestrian crossings and human control.png', p15, path = here('pngs'), width = 7, height = 7)

```

```{r}
p16 <- ggplot(data = casualty_young_fine_v %>% 
         filter(pedestrian_crossing_physical_facilities == 'No physical crossing facilities within 50 metres'),
       aes(x = vehicle_type)) +
  geom_bar(position="dodge", fill = 'blue') +
#  facet_wrap(vars(casualty_type)) +
  labs(title = 'Reported road casualties by road user type in Bradford, 2020',
       subtitle = 'Young people involved in accident classes with 200 or more casualties in 2020',
       x = 'Vehicle type',
       y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale

p16

ggsave('Road casualties by vehicle type, in the presence of pedestrian crossings and human control.pdf', p16, path = here('plots'), width = 7, height = 7)

ggsave('Road casualties by vehicle type, in the presence of pedestrian crossings and human control.png', p16, path = here('pngs'), width = 7, height = 7)

```


```{r}
p17<- ggplot(data = casualty_young_fine_v %>% 
         filter(),
       aes(x = fct_infreq(casualty_imd_decile))) +
  geom_bar(position="dodge", fill = 'blue') +
  # facet_wrap(vars(casualty_type)) +
  labs(title = 'Reported road casualties by IMD in Bradford, 2018-2020',
       subtitle = 'Young people involved in accident classes with 200 or more casualties in 2020',
       x = 'Speed limits',
       y = 'Number of accidents'
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale


p17

ggsave('Road casualties by vehicle type, in the presence of pedestrian crossings and human control.pdf', p16, path = here('plots'), width = 7, height = 7)

ggsave('Road casualties by vehicle type, in the presence of pedestrian crossings and human control.png', p16, path = here('pngs'), width = 7, height = 7)

```


```{r}
table(casualty_young_fine_v$speed_limit)  

```


```{r}
#ggsave(path = here("study1", "figs", "bar_admissions_gender_dept.png"), 
#	p, width = 7, height = 7)
```
